---
title: "Tobacco Validation Report"
author: "Xuxiang Wang", "A. Jerrod Anzalone"
date: "2025-11-27"
output: html_document
---

```{r setup, include=FALSE}
# This chunk runs automatically but is hidden in the final report.
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(gtsummary)
library(ggplot2)
# Load all the pre-computed results from the analysis script
load("../validation_results.RData")
```

### Part 1: Smoking Status Agreement and Performance

**Descriptive Statistics by Survey Smoking Status (Table 1)**

```{r}
descriptive_status
```

**Agreement between EHR-derived phenotype and self-reported survey data (Table 2)**

```{r}
table2_agreement
```

**Number of participants with selected tobacco-related indicators in wide-format EHR data (Figure 3)**

```{r}
plot_participants
```

**Number of EHR records with selected tobacco-related indicators in long-format EHR data (Figure 4)**

```{r}
plot_records
```


**Cohen's Kappa for Smoking Status**

```{r}
kappa_table_status
```

**Performance Metrics for Smoking Status (Sensitivity, Specificity, etc.)**

```{r}
performance_table_status
```

### Part 2: Smoking Intensity Agreement and Performance

**Descriptive Statistics by Survey Smoking Intensity**

```{r}
descriptive_intensity
```

**Weighted Kappa for Ordinal Smoking Intensity**

```{r}
kappa_intensity
```

**Full Confusion Matrix for Smoking Intensity**

```{r}
conf_matrix_intensity_full
```

**Per-Class, Macro, and Micro Performance for Smoking Intensity**

```{r}
performance_intensity_all
```

### Part 3: Cigarettes per Day (Continuous) Agreement

**ICC and Agreement Rate for Cigarettes per Day**

```{r}
icc_tbl_cigs
```

**Error Metrics (Bias, MAE, RMSE, Correlation) for Cigarettes per Day**

```{r}
err_tbl_cigs
```

**Bland-Altman Plot for Cigarettes per Day**

```{r}
ba_plot_cigs
```

### Part 4: Smoking Years (Continuous) Agreement

**ICC and Agreement Rate for Smoking Years**

```{r}
icc_tbl_years
```

**Error Metrics (Bias, MAE, RMSE, Correlation) for Smoking Years**

```{r}
err_tbl_years
```

**Bland-Altman Plot for Smoking Years**

```{r}
ba_plot_years
```

### Supplemental Tables: EHR Data Summaries

```{r supplemental_tables, echo=FALSE}
# This code is from the end of the original Full_Validation_Pipeline script
varlist_long <- c(
  "person_id", "concept_id", "current_smoker","former_smoker","non_smoker",
  "is_current_flag", "is_former_flag", "is_non_flag", "current_12_month","former_12_month",
  "cessation_counseling","medication_use","cessation_attempt",
  "e_cigarette_user","cigar_user","hookah_user","smokeless_user","other_type_user","cigaretee_user",
  "packs_per_day","cigarettes_per_day","intensity_daily","duration_years"
)

EHR_summary_list_long <- list()
for (v in varlist_long) {
  if (v %in% c("intensity_daily")) {
    df_sub <- long_tobacco_EHR_data %>% filter(!is.na(.data[[v]]))
  } else {
    df_sub <- long_tobacco_EHR_data %>% filter(!is.na(.data[[v]]) & .data[[v]] != 0)
  }
  EHR_summary_list_long[[v]] <- df_sub %>%
    summarise(variable = v, n_records = n(), n_persons = n_distinct(person_id))
}
sup_table_1 <- bind_rows(EHR_summary_list_long)

varlist_wide <- c(
  "ever_smoker","non_flag",
  "cessation_counseling","medication_use","cessation_attempt",
  "latest_is_current_flag","latest_is_former_flag","latest_is_non_flag",
  "latest_current_12_month","latest_former_12_month",
  "latest_cessation_counseling","latest_medication_use","latest_cessation_attempt"
)

EHR_summary_list_wide <- list()
for (v in varlist_wide) {
    df_sub <- wide_tobacco_EHR_data %>% filter(!is.na(.data[[v]]) & .data[[v]] != 0)
    EHR_summary_list_wide[[v]] <- df_sub %>%
    summarise(variable = v, n_persons = n_distinct(person_id))
}
sup_table_2 <- bind_rows(EHR_summary_list_wide)
```

**Supplemental Table 1 (Summary of Long EHR Data)**

```{r}
sup_table_1
```

**Supplemental Table 2 (Summary of Wide EHR Data)**

```{r}
sup_table_2
```

```{r, include=FALSE}
rm(descriptive_status, kappa_table_status, performance_table_status,
   descriptive_intensity, kappa_intensity, performance_intensity_all,
   icc_tbl_cigs, err_tbl_cigs, ba_plot_cigs,
   icc_tbl_years, err_tbl_years, ba_plot_years,
   sup_table_1, sup_table_2,
   EHR_summary_list_long, EHR_summary_list_wide, df_sub
)
```