---
title: "Tobacco Phenotype Validation — Results"
author: "Xuxiang Wang, A. Jerrod Anzalone"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    toc_depth: 2
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(readr)
library(gt)
library(knitr)

# Locate Results/ (knit dir differs from where results were written).
find_results <- function() {
  cands <- c("AoU-OMOP-Tobacco-Phenotyping-main/Results",
             "Results", "../Results", "../../Results",
             "../../AoU-OMOP-Tobacco-Phenotyping-main/Results")
  hit <- cands[dir.exists(cands)]
  if (length(hit) == 0) {
    stop("Results/ folder not found. Set RESULTS_DIR manually in the setup chunk.")
  }
  normalizePath(hit[1])
}
RESULTS_DIR <- find_results()

tbl_path <- function(name) file.path(RESULTS_DIR, "tables",  paste0(name, ".csv"))
fig_path <- function(name) file.path(RESULTS_DIR, "figures", paste0(name, ".png"))

# Read a result CSV and render with gt; missing files degrade gracefully.
show_tbl <- function(name, note = NULL) {
  p <- tbl_path(name)
  if (!file.exists(p)) {
    return(htmltools::HTML(sprintf("<em>Not available: %s</em>", basename(p))))
  }
  df <- readr::read_csv(p, show_col_types = FALSE)
  g <- gt::gt(df) |>
    gt::opt_row_striping() |>
    gt::tab_options(table.font.size = gt::px(13),
                    data_row.padding = gt::px(4),
                    table.width = gt::pct(100))
  if (!is.null(note)) g <- gt::tab_source_note(g, gt::md(note))
  g
}

# Show a figure PNG if present.
show_fig <- function(name) {
  p <- fig_path(name)
  if (!file.exists(p)) {
    return(htmltools::HTML(sprintf("<em>Figure not available: %s</em>", basename(p))))
  }
  knitr::include_graphics(p)
}
```

This report renders directly from the CSV tables and PNG figures written to
`Results/` by the analysis step, so every number here is reproducible from that
folder.

# Cohort

## Figure 2 — Cohort selection flow

```{r}
show_fig("figure2_cohort_flow")
```

```{r}
show_tbl("cohort_flow_counts")
```

# Part 1 — Smoking status

## Table 1 — Descriptive statistics by survey smoking status

```{r}
show_tbl("descriptive_status")
```

## Table 2 — Agreement between EHR phenotype and survey

```{r}
show_tbl("table2_agreement")
```

## Cohen's kappa — smoking status

```{r}
show_tbl("kappa_status")
```

## Performance — smoking status (sensitivity / specificity / PPV / NPV)

```{r}
show_tbl("performance_status")
```

## Figure 3 — Participants by tobacco indicator (patient level)

```{r}
show_fig("participants_by_indicator")
```

## Figure 4 — Records by tobacco indicator (event level)

```{r}
show_fig("records_by_indicator")
```

# Part 2 — Smoking intensity

## Descriptive statistics by survey intensity

```{r}
show_tbl("descriptive_intensity")
```

## Weighted kappa — intensity

```{r}
show_tbl("kappa_intensity")
```

## Confusion matrix — intensity

```{r}
show_tbl("confusion_matrix_intensity")
```

## Performance — intensity

```{r}
show_tbl("performance_intensity")
```

# Part 3 — Cigarettes per day

## Intraclass correlation

```{r}
show_tbl("icc_cigarettes_daily")
```

## Error metrics

```{r}
show_tbl("error_cigarettes_daily")
```

## Figure 5A — Bland–Altman, cigarettes per day

```{r}
show_fig("bland_altman_cigarettes")
```

# Part 4 — Smoking duration (years)

## Intraclass correlation

```{r}
show_tbl("icc_smoking_years")
```

## Error metrics

```{r}
show_tbl("error_smoking_years")
```

## Figure 5B — Bland–Altman, smoking years

```{r}
show_fig("bland_altman_smoking_years")
```

# Sensitivity analyses

## Supplemental Table 3A — Performance under varying recency (lookback) windows

```{r}
show_tbl("sensitivity_3A_recency_windows",
         note = "Recency lookback in {3, 6, 12, 24} months; pairing fixed at &plusmn;365 days.")
```

## Supplemental Table 3B — Performance stratified by survey&ndash;event time lag

```{r}
show_tbl("sensitivity_3B_timelag_strata",
         note = "Primary phenotype (12-month recency); pairing to &plusmn;730 days, stratified by absolute lag.")
```

# Supplemental — EHR summaries

## Table 1A — Patient-level variable counts

```{r}
show_tbl("supplemental_1A_patient_variables")
```

## Table 1B — Patient-level concept frequencies

```{r}
show_tbl("supplemental_1B_patient_concepts")
```

## Table 2A — Event-level variable counts

```{r}
show_tbl("supplemental_2A_event_variables")
```

## Table 2B — Event-level concept frequencies

```{r}
show_tbl("supplemental_2B_event_concepts")
```

```{r}
show_tbl("supplemental_selection_bias_comparison")
```

